\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-05-29 doc/paper.tex
remove example models E6SSM and NMSSM from the appendix

\appendix
\section{Examples}
\subsection{NMSSM}
Detailed steps to create NMSSM model file.
\begin{enumerate}
\item Create new directory \code{mkdir myNMSSM}
\item Move into this new directory \code{cd myNMSSM}
\item copy the MSSM model file in \code{cp /path/to/SARAH/Models/MSSM/MSSM.m  myNMSSM.m}
\item Edit myNMSSM.m as follows
\begin{enumerate}
\item  Change Model name, author and date
\item Add singlet superfield to SuperFields
\begin{lstlisting}
SuperFields[[8]] = {s, 1, sR,    0, 1,  1, RpM};
\end{lstlisting}
\item Modify superpotential
\begin{lstlisting}
SuperPotential = Yu u.q.Hu - Yd d.q.Hd - Ye e.l.Hd + \[Lambda] s.Hu.Hd +  \[Kappa]/3 s.s.s ;
\end{lstlisting}
\item Add new VEV
\begin{lstlisting} 
  DEFINITION[EWSB][VEVs]= 
  { {SHd0, {vd, 1/Sqrt[2]}, {sigmad, \[ImaginaryI]/Sqrt[2]},{phid,1/Sqrt[2]}},
    {SHu0, {vu, 1/Sqrt[2]}, {sigmau, \[ImaginaryI]/Sqrt[2]},{phiu,1/Sqrt[2]}},
    {SsR,  {vs, 1/Sqrt[2]}, {sigmaS, \[ImaginaryI]/Sqrt[2]},{phiS,1/Sqrt[2]}}
};
\end{lstlisting}
\item  Add the singlino to the list of gauge eigenstate dirac fermions
\begin{lstlisting}
  DEFINITION[GaugeES][DiracSpinors]={
 .............
  FS -> {FsR,conj[FsR]}   
};
\end{lstlisting}
\item Edit mass eigenstates to give how the gauge eigenstates mix into mass eigenstates.  Here we add to the cp even higgs a real singlet, to the CP odd the imaginary part of the singlet and finally to the neutralinos we add the singlino.

\begin{lstlisting}
 DEFINITION[EWSB][MatterSector]= 
{   ................
    ................
     {{phid, phiu, phiS}, {hh, ZH}},      
     {{sigmad, sigmau, sigmaS}, {Ah, ZA}},
    ................
     {{fB, fW0, FHd0, FHu0, FsR}, {L0, ZN}},
     ...............
};

\end{lstlisting}
\end{enumerate}
\item Now we copy over particles.m \code{cp /path/to/SARAH/Models/MSSM/particles.m  ./}
\item Edit particles.m as follows
\begin{enumerate}  
\item First we add to particleDefinitions[GaugeES]:
\begin{lstlisting}
 ParticleDefinitions[GaugeES] = {
      {SsR, { Description -> "Singlet"}},        
      {FS,   { Description -> "Singlino" }},    
      ............
      };
\end{lstlisting}
Note that we are brief in information here becasue there is already some information contained in the default particles.m file in the model directory:
\begin{lstlisting} 
{{ Description -> "Singlet", 
                 PDG -> {0},
                 PDG.IX -> 0,
                 Width -> 0, 
                 Mass -> Automatic,
                 FeynArtsNr -> 98,
                 LaTeX -> "S",
                 OutputName -> "s"

    }},

(* -------- Weyl Spinor  ------ *)

  {{ Description -> "Weyl Spinor of Singlino", 
                 PDG -> 99,
                 Width -> 0, 
                 Mass -> Automatic,
                 FeynArtsNr -> 9,
                 LaTeX -> "\\tilde{S}",
                 OutputName -> "s"

    }},

\end{lstlisting}
 \item Now we edit ParticleDefinitions[EWSB].  Again most states are defined in brief form by a name because they use the default information in the particles.m in the models directory.  But here we need to over write some of the default options for the cp even and cp odd higgs states: 
\begin{lstlisting}
 {hh   ,  {  Description -> "Higgs", 
                 PDG -> {25, 35,45},
                 PDG.IX ->{101000001,101000002,101000003}
 }}, 
       {Ah   ,  {    Description -> "Pseudo-Scalar Higgs",
                 PDG -> {0, 36, 46},
                 PDG.IX ->{0,102000001,102000002  } }},

\end{lstlisting}
\item To the WeylFermionAndIndermediate we add the Weyl singlino and the real scalar singlet field and imaginary scalar singlet field:  
 \begin{lstlisting}
  {sigmaS,      {Description -> "Scalar Singlet" }}  ,
  {phiS,      { Description -> "Pseudo Scalar Singlet"}},
  {FsR,   { Description -> "Weyl Spinor of Singlino"}},
 \end{lstlisting}

Although the WeylFermionAndIndermediate also contains, 
\begin{lstlisting}
{SHd,  { Description -> "Down-Higgs"}},
       {SHu,  { Description -> "Up-Higgs"}},
\end{lstlisting}
\noindent we do not add a singlet version becasue for the singlet there is no distinction between these and the individual SU(2) components, as there are for Hu and Hd.
\end{enumerate} 
\item Now we copy over parameters.m \code{cp /path/to/SARAH/Models/MSSM/parameters.m  ./}
\item Add the new parameters lambda, kappa, Alambda and Akappa to the list.
\begin{lstlisting}
{\[Kappa],   {Description -> "Singlet Self-Interaction"}},              
{T[\[Kappa]],  { Description -> "Softbreaking Singlet Self-Interaction" }}, 
{\[Lambda],   { Description -> "Singlet-Higgs-Interaction"   }},
{T[\[Lambda]],  {Description -> "Softbreaking Singlet-Higgs-Interaction"}}, 
{ms2,       { Description -> "Softbreaking Singlet Mass" }},
{vS,        { Description -> "Singlet-VEV"}}       
\end{lstlisting}
Once again this is simplified from the general case becasue there are default versions containedin parameters.m in the Models directory:
\begin{lstlisting}
{{  Description -> "Singlet Self-Interaction",
              LaTeX -> "\\kappa",
             Real ->False,
             Dependence -> None, 
             Value -> None, 
             LesHouches -> {NMSSMRUN,2},
             OutputName-> kap         }},

{{ Description -> "Softbreaking Singlet Self-Interaction",
               LaTeX -> "T_{\\kappa}",
              Real -> False,
             Dependence -> None, 
             Value -> None, 
             LesHouches ->{NMSSMRUN,4},
             OutputName-> Tk         }},

 {{ Description -> "Singlet-Higgs-Interaction",
             LaTeX -> "\\lambda",
             Real -> False,
	     Dependence -> None, 
             Value -> None, 
             LesHouches -> {NMSSMRUN,1},
             OutputName-> lam          }},

{{Description -> "Softbreaking Singlet-Higgs-Interaction",
                LaTeX -> "T_{\\lambda}",
              Real -> False,
             Dependence -> None, 
             Value -> None, 
             LesHouches ->{NMSSMRUN,3},
             OutputName-> Tlam         }},    
{{ Description -> "Softbreaking Singlet Mass", 
              LaTeX -> "m_S^2",
             DependenceNum ->  None, 
             LesHouches -> {NMSSMRUN,10},
             OutputName-> ms2 }},

{{ Description -> "Singlet-VEV", 
			 LaTeX -> "v_s",
             Value -> None, 
             LesHouches -> {NMSSMRUN,5},
             OutputName-> vS         }},
\end{lstlisting}
\item Navigate to Flexible SUSY directory  \code{cd /path/to/FlexibleSUSY/}
\item Create Z3NMSSM directory \code{./createmodel --models=Z3NMSSM:myNMSSM}
\item Configure with the new model \code{./configure --with-models=Z3NMSSM}
\item The default FlexibleSUSY model file for the MSSM will now be in the models/NMSSM2 directory.  Edit this for the NMSSM as follows:
  \begin{enumerate} 
    \item Add new input parameter $\lambda$ to EXTPAR and remove the sign of $\mu$ from MINPAR.
      \begin{lstlisting}    
        MINPAR = { {1, m0},
           {2, m12},
           {3, TanBeta},
           {5, Azero} };

  EXTPAR = { {61, LambdaInput} };
      \end{lstlisting}
    \item Name the parameters you wish to be output by the EWSB conditions.  Note you should use the names given for them in parameters.m.    A common choice in the NMSSM is:
\begin{lstlisting}
EWSBOutputParameters = {vS, \[Kappa], ms2};
\end{lstlisting}
which will work for the semi-constrained NMSSM.
\item choose the Highscale and Highscale first guess, if the default is different from what you wish to use.  
\begin{lstlisting} 
 HighScale = g1 == g2;
 HighScaleFirstGuess = 2.0 10^16;
\end{lstlisting}
\item Add how the parameters found in parameters.m that are to be set at the highscale:
\begin{lstlisting} 
HighScaleInput={
   {T[Ye], Azero*Ye},
   {T[Yd], Azero*Yd},
   {T[Yu], Azero*Yu},
   {mHd2, m0^2},
   {mHu2, m0^2},
   {mq2, UNITMATRIX[3] m0^2},
   {ml2, UNITMATRIX[3] m0^2},
   {md2, UNITMATRIX[3] m0^2},
   {mu2, UNITMATRIX[3] m0^2},
   {me2, UNITMATRIX[3] m0^2},
   {MassB, m12},
   {MassWB,m12},
   {MassG,m12},
{\[Lambda], LambdaInput},
   {T[\[Kappa]], Azero \[Kappa]},
   {T[\[Lambda]], Azero LambdaInput}
};
\end{lstlisting}
\item set lowscale boundary condititions if you want something different from the default.  The defaults are:
\begin{lstlisting} 
 LowScale = SM[MZ];
 LowScaleFirstGuess = SM[MZ];
\end{lstlisting}
\item Change the LowscaleInput if necessary.  In this example we keep the same choice as in the MSSM
  \begin{lstlisting} 
 LowScaleInput={
   {vd, 2 MZDRbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] Cos[ArcTan[TanBeta]]},
   {vu, 2 MZDRbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] Sin[ArcTan[TanBeta]]}
};
  \end{lstlisting}

Finally we set the initial guesses for the paremeters.  In this example we overwtite the MSSM version replacing it with:
 \begin{lstlisting} 
InitialGuessAtLowScale = {
   {vd, SM[vev] Cos[ArcTan[TanBeta]]},
   {vu, SM[vev] Sin[ArcTan[TanBeta]]},
   {\[Lambda], LambdaInput},
   {\[Kappa], 0.1},
   {vS, 1000},
   {ms2, SM[MZ]^2}
};

InitialGuessAtHighScale = {};
 \end{lstlisting}

The code can now be created and also compiled (if you didn't select --disable-compile option when you ran the configure script) by typing \code{make -j2}.

\end{enumerate}   
\subsection{\ESSM}
Detailed steps to create \ESSM model file.
\begin{enumerate}
\item Create new directory \code{mkdir myE6SSM}
\item Move into this new directory \code{cd myE6SSM}
\item copy the MSSM model file in \code{cp /path/to/SARAH/Models/MSSM/MSSM.m  myE6SSM.m}
\item Edit myE6SSM.m as follows
\begin{enumerate}
\item  Change Model name, author and date
\item Add new Vector superfield,
 \begin{lstlisting} 
   Gauge[[4]]={Bp,  U[1], Ncharge, g1p, False, RpM};
 \end{lstlisting}
\item Modify MSSM superfield definitions to include new charges.
\begin{lstlisting}
SuperFields[[1]] = {q, 3, {uL,  dL},     1/6, 2, 3, 1, RpM};  
SuperFields[[2]] = {l, 3, {vL,  eL},    -1/2, 2, 1, 2, RpM};
SuperFields[[3]] = {Hd,1, {Hd0, Hdm},   -1/2, 2, 1, -3, RpP};
SuperFields[[4]] = {Hu,1, {Hup, Hu0},    1/2, 2, 1, -2, RpP};
SuperFields[[5]] = {d, 3, conj[dR],    1/3, 1, -3, 2, RpM};
SuperFields[[6]] = {u, 3, conj[uR],   -2/3, 1, -3, 1, RpM};
SuperFields[[7]] = {e, 3, conj[eR],      1, 1,  1, 1, RpM};
\end{lstlisting}
Note that in the \ESSM model the charges are fixed to be those of the $U(1)_N$ as shown in e.g. \cite{King:2005jy}.  However here one can leave the charges unassigned vby using parameters in place of the fixed charges.  Here we fix the charges to simplify the C++ code.

\item Add Singlet Higgs superfield to SuperField
\begin{lstlisting}
SuperFields[[8]] = {s, 1, sR,     0, 1,  1, 5, RpP};
\end{lstlisting}
\item Add inert Higgs-like superfields 
\begin{lstlisting}
SuperFields[[9]] = {H1I, 2, {H1I0, H1Im},  -1/2, 2, 1, -3, RpP};
SuperFields[[10]] = {H2I, 2, {H2Ip, H2I0},   1/2, 2, 1, -2, RpP};
SuperFields[[11]] = {sI, 2, sIR,    0, 1,  1, 5, RpM};
\end{lstlisting}
\item Add exotics to SuperFields
\begin{lstlisting}
SuperFields[[12]] = {Dx, 3, DxL,  -1/3, 1, 3, -2, RpP};
SuperFields[[13]] = {Dxbar, 3, conj[DxbarR],  1/3, 1, -3, -3, RpP};
\end{lstlisting}
\item Add additional survival $SU(2)$ multiplets to SuperFields
\begin{lstlisting}
SuperFields[[14]] = {Hp, 1, {Hpd0, Hpdm},  1/2, 2,  1, 2, RpP};
SuperFields[[15]] = {Hpbar, 1, {Hpup, Hpu0}, -1/2, 2,  1, -2, RpP};
\end{lstlisting}
\item Modify superpotential
\begin{lstlisting}
SuperPotential = Yu u.q.Hu - Yd d.q.Hd - Ye e.l.Hd + \[Lambda] s.Hu.Hd +  \[Lambda]12 s.H2I.H1I + \[Kappa] s.Dx.Dxbar + \[Mu]Pr Hp.Hpbar;
\end{lstlisting}
\item Define new gauge eigenstate Dirac spinors
\begin{lstlisting}
  DEFINITION[GaugeES][DiracSpinors]={
    .............
    .............
    FS -> {FsR,conj[FsR]},
    FBp -> {fBp,conj[fBp]},
    HI0 -> {FH1I0, conj[FH2I0]},
    HIC -> {FH1Im, conj[FH2Ip]},
    FSI -> {FsIR, conj[FsIR]},
    FDx1 -> {FDxL, 0},
    FDx2 -> {0, FDxbarR}, 
    Hp0 -> {FHpd0, conj[FHpu0]},
    HpC -> {FHpdm, conj[FHpup]}
  };
\end{lstlisting}
\item Add new VEV
\begin{lstlisting} 
  DEFINITION[EWSB][VEVs]= 
  { {SHd0, {vd, 1/Sqrt[2]}, {sigmad, \[ImaginaryI]/Sqrt[2]},{phid,1/Sqrt[2]}},
    {SHu0, {vu, 1/Sqrt[2]}, {sigmau, \[ImaginaryI]/Sqrt[2]},{phiu,1/Sqrt[2]}},
    {SsR,  {vs, 1/Sqrt[2]}, {sigmaS, \[ImaginaryI]/Sqrt[2]},{phiS,1/Sqrt[2]}}
};
\end{lstlisting}
\item Add new gauge boson  mass eigenstate
\begin{lstlisting} 
  DEFINITION[EWSB][GaugeSector] =
  { 
    {{VB,VWB[3],VBp},{VP,VZ,VZp},ZZ},
    {{VWB[1],VWB[2]},{VWm,conj[VWm]},ZW},
    {{fWB[1],fWB[2],fWB[3]},{fWm,fWp,fW0},ZfW}
  }; 
\end{lstlisting}
\item Edit mixings for Higgs and neutralino states and add additional states that mix.
\begin{lstlisting}
DEFINITION[EWSB][MatterSector]= 
{ ...............
  ...............
{{phid, phiu, phiS}, {hh, ZH}},
{{sigmad, sigmau, sigmaS}, {Ah, ZA}},
  ...............
{{fB, fW0, FHd0, FHu0, FsR, fBp}, {L0, ZN}},
 ................ 
 {{SH1I0,conj[SH2I0]},{SHI0,UHI0}},
 {{SH1Im,conj[SH2Ip]},{SHIp,UHIp}},
 {{SHpd0,conj[SHpu0]},{SHp0,UHp0}},
 {{SHpdm,conj[SHpup]},{SHpp,UHpp}},
 {{FHpd0,FHpu0},{L0p,ZNp}},
 {{FH1I0,FH2I0},{L0I,ZNI}}
\end{lstlisting}
\end{enumerate}

\section{Numeric tests}

>>>>

\appendix
\section{Numeric tests}
